---
title: "Report"
author: "Gwyneth Wei (gw2442); Emma Sexton (els2250); Olivia Wang (hw2852); Mayuri Albal (ma4197) <br>"
date: "10 December 2022"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### In-depth Analysis of Emma Sexton's New York City 2022 Marathon Training

<br/>  

#### Motivation: Provide an overview of the project goals and motivation.

<br/>  

#### Related Work: Anything that inspired you, such as a paper, a web site, or something we discussed in class.

<br/>  

#### Initial Questions: What questions are you trying to answer? How did these questions evolve over the course of the project? What new questions did you consider in the course of your analysis?

<br/>  

#### Data: Source, scraping method, cleaning, etc.
Data was collected using the Garmin Forerunner 245 activity watch from August 17, 2020 to November 6, 2022. Collected data was uploaded to Strava.com after each activity. After granting permission, personal data was downloaded from Strava.com. Each activity logged by the Garmin watch was downloaded as a separate FIT file; therefore, using the `FITfileR` package, data was imported into R. 

##### *Importing Data*

First, we imported the summary data of all of Emma's training runs from July 23, 2020...

```{r, load packages}
library(tidyverse)
library(FITfileR)
library(leaflet)
```



```{r}
training_summary <- read_csv("activities/activities.csv") %>% 
  janitor::clean_names() %>%
  filter(activity_type == "Run") %>%
  select(activity_id, activity_date, activity_name, activity_description, elapsed_time_6, distance_7, max_heart_rate_8, relative_effort_9, max_speed, average_speed, elevation_gain, elevation_loss, max_grade, average_grade, max_cadence, average_cadence, average_heart_rate, calories, weather_temperature, dewpoint, humidity, wind_speed) %>%
  filter (activity_id >= 6910869137) %>%
  separate(activity_date, c("month_date", "year", "time"), sep = ", ") %>%
  mutate(
    date = str_c(month_date, year, sep = " "),
    date = as.Date(date, format = "%b%d%Y")
  ) %>%
  select(-month_date, -year) %>%
  mutate(
    elapsed_time_min = elapsed_time_6 / 60,
    distance_km = distance_7,
    max_heart_rate = max_heart_rate_8,
    relative_effort = relative_effort_9
  ) %>%
  select(-elapsed_time_6, -distance_7, -max_heart_rate_8, -relative_effort_9)

activity_summaries =
  training_summary %>%
  select(activity_id, date, time, activity_name, activity_description)

tidy_training = 
  training_summary %>%
  select(-activity_name, -activity_description) %>%
  select(activity_id, date, time, distance_km, elapsed_time_min, max_speed, average_speed, max_heart_rate, average_heart_rate, relative_effort, everything())
```

Since each Garmin activity was downloaded as an individual FIT file, a function was created to iterate across the multiple activities and compile a single data frame. As shown below, the function was created to read each activity FIT file into R and pull out the records (which included data on latitude and longitude, distance, speed, heart rate, and cadence), and then each second of activity was merged into a single data frame. This function was applied to a list of all file names. A `for` loop was used to generate a list of data frames, and then rows were merged using `bind_rows` to create the final data frame, `marathon_df`. 

##### *Data Cleaning*
After importing, the data frame was then cleaned to include activities from the start of marathon training (March 31, 2022) to the day of the New York City Marathon (November 6, 2022). Within the 'activities' data set, there were 335 observations and 84 variables. The data was filtered to only include observations from 'activity_type' that were classified as 'Run'. From this, 22 variables were selected to be included in this data set: activity_id, activity_date, activity_name, activity_description, elapsed_time_6, distance_7, max_heart_rate_8, relative_effort_9, max_speed, average_speed, elevation_gain, elevation_loss, max_grade, average_grade, max_cadence, average_cadence, average_heart_rate, calories, weather_temperature, dewpoint, humidity,and wind_speed. 

Due to Emma's marathon training beginning March 31st, 2022, the 'activity_id' was used in order to filter any training prior to this. Following this, the variable 'activity_date' was separated into two variables:the date on which the run occurred ('date') as well as the time when the run began ('time'). The 'elapsed_time_6' variable was converted from seconds to minutes ('elapsed_time_min'). In addition to this, the variables names for distance_7, max_heart_rate_8, and relative_effort_9 were changed to distance_km, max_heart_rate, and relative_effort.

A new data frame, 'activities_summaries' was created in order to include the written accounts from each of Emma's runs, including variables 'activity_name' and 'activity_description'.

The final cleaned data set, 'tidy_training', included 110 observations and 21 variables. Variables included were:activity_id, date, time, distance_km, elapsed_time_min, max_speed, average_speed, max_heart_rate, average_heart_rate, relative_effort, elevation_gain, elevation_loss, max_grade, average_grade, max_cadence, averag_cadence, calories, weather_temperature, dewpoint, humidity, and wind_speed. 

```{r, cache = TRUE}
# data_import <- function(list_data) {
# 
#   garmin_data <- readFitFile(list_data)
# 
#   garmin_activities_df <- records(garmin_data) %>%
#     bind_rows() %>%
#     arrange(timestamp)
# 
#   garmin_activities_df
# 
# }
# 
# garmin_fitfiles <- "activities/fit_files/garmin_data_"
# 
# list_of_fitfiles <- str_c(garmin_fitfiles, c("1":"329"), ".fit")
# 
# n = 329
# datalist = list()
# datalist = vector("list", length = n)
# 
# for (i in 1:n) {
#   datalist[[i]] <- data_import(list_of_fitfiles[i])
# }
# 
# marathon_df <- bind_rows(datalist)
```

##### *Data Cleaning*
After importing, the data frame was then cleaned to include activities from the start of marathon training (March 31, 2022) to the day of the New York City Marathon (November 6, 2022). Data were...

Other decisions we need to make:  

[ ] Need to decide if we want to keep each observation as a single second for each activity  

[ ] Need to decide if we want to focus solely on training runs or include walks and other activities (elliptical or biking)

<br/>  

#### Exploratory Analysis: Visualizations, summaries, and exploratory statistical analyses. Justify the steps you took, and show any major changes to your ideas.

* Mileage over time (Intro)
* Mapping running trails (plotly/gis(?))
  Map marathon route and the average speed per mile
* Temperature vs average speed/average heart rate/relative effort
* Elevation vs average speed/relative effort 
* Dashboard with all our plots

<br/>  

#### Additional Analysis: If you undertake formal statistical analyses, describe these in detail

<br/>  

#### Discussion: What were your findings? Are they what you expect? What insights into the data can you make?


*Map of Marathon Route with Averages*
NOTE: Install Leaflet Prior to Running

Loading and Cleaning the Data
```{r}
marathon_day <- read_csv("activities/marathon_day.csv")

marathon_day=
marathon_day %>% 
mutate(
  distance_m= distance*0.000621371
) %>% 
mutate(group = case_when(
    between(distance_m, 0,0.9999) ~ "Start",
    between(distance_m, 1, 1.9999) ~ "1",
    between(distance_m, 2, 2.9999) ~ "2",
    between(distance_m, 3, 3.9999) ~ "3",
    between(distance_m, 4, 4.9999) ~ "4",
    between(distance_m, 5, 5.9999) ~ "5",
    between(distance_m, 6, 6.9999) ~ "6",
    between(distance_m, 7, 7.9999) ~ "7",
    between(distance_m, 8, 8.9999) ~ "8",
    between(distance_m, 9, 9.9999) ~ "9",
    between(distance_m, 10, 10.9999) ~ "10",
    between(distance_m, 11, 11.9999) ~ "11",
    between(distance_m, 12, 12.9999) ~ "12",
    between(distance_m, 13, 13.9999) ~ "13",
    between(distance_m, 14, 14.9999) ~ "14",
    between(distance_m, 15, 15.9999) ~ "15",
    between(distance_m, 16, 2.9999) ~ "16",
    between(distance_m, 17, 17.9999) ~ "17",
    between(distance_m, 18, 18.9999) ~ "18",
    between(distance_m, 19, 19.9999) ~ "19",
    between(distance_m, 20, 20.9999) ~ "20",
    between(distance_m, 21, 21.9999) ~ "21",
    between(distance_m, 22, 22.9999) ~ "22",
    between(distance_m, 23, 23.9999) ~ "23",
    between(distance_m, 24, 24.9999) ~ "24",
    between(distance_m, 25, 25.9999) ~ "25",
    between(distance_m, 26, 26.9999) ~ "26-Finish")) 
```

Recode
```{r}
marathon_geo=
  marathon_day %>% 
  mutate(
    latitude = position_lat,
    longitude= position_long) %>% 
  select(-position_lat, -position_long)
```

```{r}
marathon_speed=
marathon_geo %>% 
  group_by(group) %>% 
  mutate(
    speed= 
  mean(enhanced_speed))
```

Aggregate Data
```{r}
agg = aggregate(marathon_geo,
                by = list(marathon_geo$group),
                FUN = mean)
```

Aggregate Leaflet Plot

```{r}
leaflet() %>% 
  addTiles() %>% 
  addCircleMarkers(data = agg,
             lng = ~longitude,
             lat = ~latitude,
             label = ~Group.1,
             radius = 7,
             color= "orange",
             stroke = TRUE, fillOpacity = 0.75,
             popup = ~paste("Mile:", Group.1,
                            "<br>Speed:", enhanced_speed,
                            "<br>HR:", heart_rate,
                             "<br>Cadence:", cadence,
                            "<br>Altitude:", enhanced_altitude)) %>% 
  addProviderTiles(providers$CartoDB.Positron)
```

